---
title: "R Notebook"
output: html_notebook
---

### Part 1 - Burglaries

You've been provided three datasets for this project:
* burglaries_2023.csv: Contains data on the aggravated burglary incidents in Davidson County. This was obtained from https://experience.arcgis.com/experience/060537e33bff45d996ca09f84a4eb0a6/.
* census.csv: Census tract level data on population and median income. This was obtained from the US Census American Community Survey.
* DC: A shapefile containing Davidson County census tracts

Perform a spatial join to determine the census tract in which each burglary occurred.

```{r}
library(sf)
library(tidyverse)
```
```{r}
burglaries_2023 <- read_csv('../data/burglaries_2023.csv')
burglaries_2023
```
```{r}
census <- read_csv('../data/census.csv')
 
  
```
```{r}
DC <- read_sf('../data/DC')
DC
```
```{r}
burglar_geo <- st_as_sf(
  burglaries_2023 |> drop_na('latitude','longitude'), 
    coords = c('longitude', 'latitude'),
  crs = st_crs(DC)
)
burglar_geo
```
```{r}
DC <- DC |> 
  mutate(
    latitude = as.numeric(INTPTLAT),
    longitude = as.numeric(INTPTLON)
    )
```

```{r}
burg_DC <- st_join(burglar_geo, DC, join=st_within, left=FALSE)
burg_DC
```
```{r}
burg_DC <- burg_DC |> 
  rename('tract' = 'TRACTCE') 
burg_DC
  
```

```{r}
burg_DC_census <- burg_DC |> 
  left_join(census, by = 'tract')
burg_DC_census
```

Perform some exploraty analysis on your prepared dataset.

Aggregate the data by census tract. **Warning:** each incident can appear multiple times if there are multiple victims, so be sure that you aren't double-counting any incidents. 

Which census tract had the highest number of burglaries? Which census tract had the highest number of burglaries per 1000 residents? 

Finally, look at the relationship between median income and number of aggravated burglaries per tract. How would you describe this relationship?

```{r}
burg_DC_census |> 
  filter(grepl('WHITE', incident_location)) |> 
  count()
```
```{r}
burg_tract <- burg_DC_census |> 
  filter(median_income>0) |> 
  group_by(tract) |> 
  distinct(incident_number, .keep_all = TRUE)
burg_tract  
```

```{r}
burg_count <- burg_tract |>
  count(tract) |> 
  arrange(desc(n))
burg_count  
```
Census tract 16000 had the highest number of burglaries.

```{r}
burg_count <- burg_count |>
  left_join(census, by = 'tract') |>
  mutate(burg_per_thousand = n * 1000/population) |> 
  arrange(desc(burg_per_thousand))

```
Tract 16000 has the highest number of burglaries per 1,000 residents.
```{r}
burg_count |> 
  st_drop_geometry() |> 
  ggplot(aes(median_income, n)) + geom_point()
```
There seems to be a negative exponential relationship between median income and number of burglaries per tract, the higher the median income the lower the number of burglaries.

